The \dealii{} Library, Version 8.5

### 2017-03-03 8.5/paper.tex
Improve text

For the new release, the geometric multigrid in \dealii{} have been thoroughly
overhauled regarding their scalability on large-scale parallel computers. To
this end, the geometric multigrid algorithm based on the fast matrix-free
kernels from \cite{KronbichlerKormann2012} have been benchmarked up to
147,456 cores. Several scalability bottlenecks have been removed, including
unnecessary inner products inside the Chebyshev smoother and
$\mathcal O(n_\text{levels})$ global communication steps during the
restriction process rather than only the single global communication step that
is necessary when going to the coarest grid. New matrix-free transfer
implementations called \texttt{MGTransferMatrixFree} were devised that replace
the matrix-based \texttt{MGTransferPrebuilt}. Besides better scalability than
the Trilinos Epetra matrices underlying the latter, the matrix-free transfer
is also a much better for high-order elements with complexity per degree of
freedom of $\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions
rather than $\mathcal O(p^d)$ for the matrix-based approach.

@@

The scalability of the improved geometric multigrid framework is shown in
Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scalability
plot in the left panel using continuous $\mathcal Q_3$ elements on 57 million
to 232 billion degrees of freedom for the Laplacian. Along each line, the same
problem size is solved with an increasing number of cores, whereas different
lines always start out at 3.5 million degrees of freedom per core. Almost
ideal scalability down to approximately 0.1 seconds can be observed also on
147k cores. The right panel of Fig.~\ref{fig:scaling_mg} shows the effect of
the aforementioned algorithmic improvements on a setup with discontinuous DG
elements, clearly improving the latency of the multigrid V-cycle. The improved
algorithms are shown in the updated step-37 tutorial program.

>>>>

For the new release, the geometric multigrid facilities in \dealii{} have been
thoroughly overhauled regarding their scalability on large-scale parallel
computers. During this process, a geometric multigrid implementation based on
the fast matrix-free kernels from \cite{KronbichlerKormann2012} has been
benchmarked up to 147,456 cores. The fast matrix-vector product revealed
several scalability bottlenecks, including unnecessary inner products inside
the Chebyshev smoother and $\mathcal O(n_\text{levels})$ global communication
steps during the restriction process rather than only the single global
communication step that is necessary when going to the coarest grid. New
matrix-free transfer implementations called \texttt{MGTransferMatrixFree} were
devised that can replace the matrix-based \texttt{MGTransferPrebuilt} class
for tensor product elements. Besides better scalability than the Trilinos
Epetra matrices underlying the latter, the matrix-free transfer is also a much
better for high-order elements with complexity per degree of freedom of
$\mathcal O(d p)$ in the polynomial degree $p$ in $d$ dimensions rather than
$\mathcal O(p^d)$ for the matrix-based approach.

@@

The scalability of the improved geometric multigrid framework is shown in
Fig.~\ref{fig:scaling_mg}, including a combined strong and weak scaling plot
in the left panel using continuous $\mathcal Q_3$ elements with 57~million to
232~billion degrees of freedom for discretizing the Laplacian. Along each
line, the same problem size is solved with an increasing number of cores,
whereas different lines are a factor of eight apart and always start at
3.5~million degrees of freedom per core. Almost ideal scalability down to
approximately 0.1~seconds can be observed also on 147k~cores. The right panel
of Fig.~\ref{fig:scaling_mg} shows the effect of the aforementioned
algorithmic improvements on a setup with discontinuous DG elements, clearly
improving the latency of the multigrid V-cycle. The improved algorithms are
shown in the updated step-37 tutorial program.
